---
title: "RNAseq Pathway Enrichment Analysis"
author: "Hao He"
date: "`r Sys.time()`"
editor_options:
  chunk_output_type: console
params:
  # project directory
  root_dir: "D:\\20260211_RNA_seq"
  # hs or mm
  species: "hs" 
  # general setting
  samplesheet: "samplesheet.csv" 
  deg_padj: 0.01
  deg_fc: 1.5
  deg_threshold: 2 # tpm threshold at least in one group for plot heatmmap
  deseq2_dir: "deseq2"
  enrichment_dir: "enrichment"
  wald_test_suffix: "_wald_result.xlsx" 
  expr_file_suffix: "_norm_expression.xlsx"
  quantification_dir: "quantification"
--- 

# Setup
```{r, include=FALSE, cache = FALSE}
# options("repos" = c(CRAN = "http://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
# options(BioC_mirror = "http://mirrors.tuna.tsinghua.edu.cn/bioconductor")
script_path <- rstudioapi::getActiveDocumentContext()$path |> dirname()
source(file.path(script_path, "RNAseq_helper.R"))

# knitr::opts_knit$set(root.dir=dirname(f_path))
setwd(params$root_dir)
knitr::opts_chunk$set(echo=TRUE, results="hold", cache=TRUE)

species <- params$species
deg_padj <- params$deg_padj
deg_fc <- params$deg_fc
deg_threshold <- params$deg_threshold
deseq2_dir <- params$deseq2_dir
wald_test_suffix <- params$wald_test_suffix
enrichment_dir <- params$enrichment_dir
expr_file_suffix <- params$expr_file_suffix
quantification_dir <- params$quantification_dir
samplesheet <- params$samplesheet

# require(OmnipathR)
# library(progeny)
# library(decoupleR)
library(tidyverse)
library(msigdbr)
library(clusterProfiler)
library(GSVA)
library(enrichplot)
library(pathview)
library(pheatmap)
library(factoextra)
library(ComplexHeatmap)
library(GseaVis)
require(smerc)
require(circlize)
#devtools::install_github('ievaKer/aPEAR')
library(aPEAR)
library(patchwork)
library(rio)
options(rio.import.class='tbl')
library(furrr)
plan(multisession, workers=8)
# plan(sequential)
```


## read sample info
```{r}
sample <- import(samplesheet)
```

## read expression
```{r}
# all deg must be within expr filtered mat
expr_file <- list.files(
  quantification_dir, expr_file_suffix, full.names = T) %>% 
  str_subset("^[^~]") %>% last() %>% print()

# convert to matrix
use_data <- "tpm" # tpm, vst, counts_norm
expr_data <- import(expr_file, which=use_data) %>%
    select(all_of(c("symbol", sample$sample))) %>%
  # filter by expr
  filter_df_by_group(
    sample[["group"]] %>% set_names(sample$sample), 
    threshold = deg_threshold) %>% 
  filter(!is.na(symbol)) %>%
  distinct(symbol, .keep_all=T)

# check dims
dim(expr_data)
expr_data[1:5,1:5]
```


# Comparison Based
## read wald result
```{r}
# get wald test result file
wald_test_file <- list.files(deseq2_dir, wald_test_suffix, full.names = T) %>% 
  str_subset("^[^~]") %>% last() %>% print()

# read multi-sheet data
comparison <- readxl::excel_sheets(wald_test_file)
cat(str_glue("{str_c(comparison, collapse='\n')}"))
wald_test <- comparison %>% set_names(comparison) %>% 
  map(\(x) import(wald_test_file, which=x) %>% 
        # order by stats
        arrange(desc(stat), pvalue) )

# get ranked genelist, for GSEA
genelist_ranked <- wald_test %>% 
  map(\(x) distinct(x, symbol, .keep_all = T) %>% 
        arrange(desc(stat)) %>% 
        select(symbol, stat) %>% deframe())

# subset sig deg
print("filter out comparison by < 20 sig genes")
wald_test_sig <- wald_test %>% 
  map(\(x) filter(
    x, abs(log2FoldChange) > log2(deg_fc), padj < deg_padj) %>% 
      {if(exists("expr_data")) filter(., symbol %in% expr_data$symbol) else .} ) %>% 
  # remove comparison which has less sig genes
  .[map_int(., nrow) > 20] %>% 
  map(\(x) pull(x, symbol))

glimpse(wald_test_sig)
```

## make geneset
```{r}
if(!exists("genesets")) {
  print("prepare term2gene geneset from msigdbr")
  # msigdbr::msigdbr_show_species()
  # msigdbr::msigdbr_collections()
  
  # msigdb <- msigdbr::msigdbr(
  #   species = if_else(species=="hs", "Homo sapiens", "Mus musculus"))
  
  ## subset genesets
  genesets <- list(
    ## H
    # Hallmark = filter(msigdb, gs_cat == "H")
    ## GO
    # GOBP = filter(msigdb, gs_subcat == "GO:BP")
    # GOMF = filter(msigdb, gs_subcat == "GO:MF"),
    # GOCC = filter(msigdb, gs_subcat == "GO:CC"),
    ## cannonical pathway
    # KEGG = filter(msigdb, gs_subcat == "CP:KEGG"),
    # PID = filter(msigdb, gs_subcat == "CP:PID"),
    # BIOCARTA = filter(msigdb, gs_subcat == "CP:BIOCARTA"),
    # WikiPathway = filter(msigdb, gs_subcat == "CP:WIKIPATHWAYS"),
    # Reactome = filter(msigdb, gs_subcat == "CP:REACTOME")
    ## TF
  ) %>% 
    # get gs_name to tidy format
    map(\(x) dplyr::select(x, gs_id=gs_exact_source, gs_name, gene_symbol) %>% 
          mutate(gs_id=ifelse(gs_id=="", gs_name, gs_id)))
    
  ## self defined geneset
  if(F) {
    yap_targets <- c(
      "CTGF","CCN2","ANKRD1","CYR61","CCN1","TOP2A","KIF14","CCNA2",
      "CDCA8","CENPF","KIF23","KIF20B","KNTC1","RRM2",
      "MCM3","SGOL1.AS1","TUBB","MYBL1","RAD18",
      "ZWILCH","SGOL1","TIMELESS","GINS1","SMC3","TK1",
      "MRE11A","MCM7","SUV39H2","GADD45B","FOSL1","CENPV",
      "RUVBL2","MYC","GLI2","AXL","ABCB1","CAT","GPATCH4",
      "LMNB2","TXN","WSB2","AREG","FOXF2","IGFBP3","RASSF2",
      "AMOTL2","NPPB","CCND1") %>%
      str_trim() %>% unique() %>% str_to_upper()
    yap_targets <- crossing(
      gs_id="HALLMARK_YAP_TARGETS",
      gs_name = "HALLMARK_YAP_TARGETS", 
      gene_symbol = yap_targets)
    # add to Hallmark
    genesets$Hallmark <- genesets$Hallmark %>% 
      bind_rows(yap_targets)
  }
  
  # to list
  genesets <- genesets %>% 
    map(\(x) list(term2gene=x[, c("gs_id", "gene_symbol")],
                  term2name=x[, c("gs_id", "gs_name")]))
}
```

## DEGs cluster
### optimal k
```{r}
mat <- expr_data[,] %>% 
  filter(symbol %in% unique(unlist(wald_test_sig))) %>%
  # filter(symbol %in% yap_targets$gene_symbol) %>%
  column_to_rownames("symbol") %>% 
  scale_mat(by="row")
mat[1:5,1:5]

if(T) {
  # optimal clusters
  elbow_plot <- mat %>% 
    # fviz_nbclust(kmeans, method="silhouette", k.max=20)
    fviz_nbclust(kmeans, method="wss", k.max=20)
  print(elbow_plot)
  k <- smerc::elbow_point(as.integer(elbow_plot$data$clusters), elbow_plot$data$y) %>% 
    .[["x"]] %>% print()
} else k <- 4 # defined by hand
```

### mark gene
```{r}
# mark gene
pathway <- c("HALLMARK_MYC_TARGETS_V1","HALLMARK_MYC_TARGETS_V2")
mark_gene <- list_flatten(genesets) %>% 
  .[str_detect(names(.), "term2gene")] %>% 
  list_rbind() %>% 
  filter(gs_id %in% pathway) %>% 
  pull(gene_symbol) %>% 
  intersect(rownames(mat))
# mark_gene <- sample(rownames(mat), 40)
mark_gene <- anno_mark(
  at = match(mark_gene, rownames(mat)), labels = mark_gene,
  labels_gp = gpar(fontsize=5), which = "row")
# mark_gene <- NULL
```

### annotation
```{r}
# sample anno
anno_col_list <- annoCol_from_sample_info(sample, split_by="group", annotate_col=NULL)

# cluster by row
anno_cl_list <- annoRow_from_cluster(mat, k)

# box trend plot by cluster
box_color <- anno_col_list$sample_split_color
plot_boxtrend <- as_tibble(mat, rownames="symbol") %>% 
  bind_cols(cluster=anno_cl_list$cl$cluster, .) %>% 
  pivot_longer(!c(cluster, symbol), names_to="sample") %>% 
  left_join(sample, by = join_by(sample)) %>% 
  split(., .[["cluster"]]) %>% 
  map(\(x) ggplot(x) + geom_boxplot(
    aes(contrast, value), 
    # fill=unique(box_color) %>% set_names(unique(names(box_color))),
    color=unique(box_color) %>% set_names(unique(names(box_color))),
    width=0.4, linewidth=0.5, outliers=F, staplewidth=0.5) +
      geom_line(
        aes(contrast, value, group=1), linewidth=1, color="red3",
        data=reframe(x, value=mean(value), .by=contrast)) +
      labs(x="", y="") +
      theme_bw(4) +
      theme(axis.text.x = element_blank(),
            axis.ticks = element_blank(),
            panel.grid = element_blank()) )
# plot_boxtrend[[1]]
anno_boxtrend <- anno_gglist(
  align_to=anno_cl_list$align_to, gglist=plot_boxtrend, 
  # panel size, gap, width, fill, color
  ggplot.panel.arg = c(2, 0.4, 4, "grey90","grey30"))

# anno_BoxTrend(
#   mat, 
#   align_to=anno_cl_list$align_to, cl=anno_cl_list$cl, 
#   sample_df=sample, group_name="group", 
#   group_color=anno_col_list$sample_split_color,
#   # panel size, gap, width, fill, color
#   ggplot.panel.arg = c(2, 0.4, 4, "grey90","grey30"))

# enrichment by cluster
enrich_hallmark <- anno_cl_list$cl$cluster %>% split(., .) %>% 
  modify(names) %>% enirch_ora_list(genesets["Hallmark"])
plot_hallmark <- enrich_hallmark %>% map(
  \(x) x@result %>% 
    mutate(.,Description=factor(
      Description, levels=rev(unique(.$Description)))) %>% 
    filter(pvalue < 0.01) %>% 
    slice_min(n=5, order_by=p.adjust) %>% 
    ggplot(aes(-log10(p.adjust), Description)) + 
    geom_col(aes(fill=-log10(p.adjust)), width=0.8, show.legend = F) +
    geom_text(aes(x=0, label=Description), #size=2,
              hjust=0, nudge_x=0.05, show.legend = F) +
    # scale_y_discrete(labels=function(x) str_wrap(x, width=40)) +
    scale_fill_gradient(low="#fee8c8", high="#ef6548") +
    theme_bw(2) + labs(x="",y="") +
    theme(axis.ticks = element_blank(),
          axis.text = element_blank(),
          panel.grid = element_blank()) )
# plot_hallmark[[1]]
anno_hallmark <- anno_gglist(
  align_to=anno_cl_list$align_to, gglist=plot_hallmark, 
  # panel size, gap, width, fill, color
  ggplot.panel.arg = c(2, 0.4, 6, "grey90","grey30"))

# add row annotation together
right_annotation = rowAnnotation(
  gene=mark_gene, cluster=anno_cl_list$anno_cl, 
  trend=anno_boxtrend, hallmark=anno_hallmark)
```

### display heatmap
```{r}
# options
if("fastcluster" %in% installed.packages()) ht_opt$fast_hclust <- TRUE
ht_opt$TITLE_PADDING = unit(c(4, 4), "points")
ht_opt$message = FALSE

# plot
pdf(str_glue("{enrichment_dir}/{Sys.Date()}_deg_all_cluster.pdf"), 
    width=16, height=10)
set.seed(42)
p <- Heatmap(
  mat, name = " ",  border = T,
  col = circlize::colorRamp2(c(-2.5, 0, 2.5), c("#2166ac", "white", "#b2182b")),
  
  cluster_columns = F,
  column_split = anno_col_list$sample_split,
  column_title = NULL,#unique(sample_split),
  # column_title_gp = gpar(fontsize=6, fill=sample_split_color),
  show_column_names = T, column_names_side = "bottom",
  column_names_gp = gpar(fontsize=4),
  column_names_rot = 45,
  top_annotation = anno_col_list$anno_sample,
  
  row_split = anno_cl_list$cl$cluster,
  row_title = "C%s",
  row_title_gp = gpar(fontsize=8),
  row_title_rot = 0,
  row_title_side = "left",
  right_annotation = right_annotation,
  show_row_names = F,
  row_names_gp = gpar(fontsize=4)) %>% 
  draw(use_raster=T)
# grid.newpage()
# pushViewport(viewport(layout = grid.layout(nr = 1, nc = length(ks))))
# for(i in seq_along(ht_list)) {
#   pushViewport(viewport(layout.pos.row = 1, layout.pos.col = i))
#   draw(ht_list[[i]], newpage = F)
#   upViewport() }
while(dev.cur() != 1) dev.off()

# save data
row_ord <- enframe(row_order(p), "cluster", "row_order") %>% 
  unnest(row_order)
ht_cluster <- mat[row_ord$row_order, unlist(column_order(p))] %>% 
  round(3) %>% 
  cbind(cluster=row_ord$cluster, .) %>% 
  as_tibble(rownames="symbol")
export(ht_cluster, str_glue(
  "{enrichment_dir}/{Sys.Date()}_deg_all_cluster.xlsx"))
```


## ORA (pairwise)
### enrichment
```{r}
enrich_result_ora <- enrich_ora_list(wald_test_sig, genesets)

# save data
export(map(enrich_result_ora, \(x) as_tibble(x@result)), 
           str_glue("{enrichment_dir}/{Sys.Date()}_enrichment_ora.xlsx"))
# save plot
n_col <- length(enrich_result_ora) #if_else( > 1, 2, 1)
map(enrich_result_ora, 
    \(x) plot_enrich_ora(x@result, facet_by="Category")) %>%
  plot_grid(plotlist = ., 
            labels = names(enrich_result_ora), 
            ncol = n_col, 
            axis = "tblr", align = "hv", 
            label_size = 8, scale = 0.9) %>% 
  ggsave(str_glue("{enrichment_dir}/{Sys.Date()}_enrichment_ora.pdf"),
          .,
          width=n_col*6, 
          height=length(genesets)*4, 
          limitsize = F)
```

### visNetwork
```{r}
pvalue_cutoff <- 1e-3
min_cluster_size <- 3
plist <- list()

for(x in names(enrich_result_ora)) {
  print(str_glue("visNetwork >>> {x}"))
  plist[[x]] <- enrich_result_ora[[x]]@result %>% 
    filter(pvalue < pvalue_cutoff) %>% 
    aPEAR::enrichmentNetwork(
      simMethod = "jaccard", # c('jaccard', 'cosine', 'cor'),
      clustMethod = "markov", # c('markov', 'hier', 'spectral'),
      clustNameMethod = "pagerank", # c('pagerank', 'hits', 'none'),
      minClusterSize = min_cluster_size, # min number of nodes in a single cluster
      colorBy = "pvalue", # pvalue, NES
      colorType = "pval", # use log transform on the colorBy column and adjust color range
      nodeSize = "Count",
      innerCutoff = 0.1, # similarity cutoff for in-cluster nodes
      outerCutoff = 0.5, # similarity cutoff for between-cluster nodes
      pCutoff = -5, # adjust p-value colouring cutoff when using pval
      drawEllipses = T,
      fontSize = 5,
      repelLabels = T,
      plotOnly = T,
      verbose = F)
}

# save plot
plot_grid(plotlist = plist, 
          labels = names(enrich_result_ora), 
          ncol = 1, 
          axis = "tblr", align = "hv", 
          label_size = 8, scale = 0.9) %>% 
  ggsave(str_glue(
    "{enrichment_dir}/{Sys.Date()}_enrichment_ora_visNetwork.pdf"),
    .,
    width=10, 
    height=length(enrich_result_ora)*10, 
    limitsize = F)
```


## GSEA (pairwise)
### enrichment
```{r}
enrich_result_gsea <- enrich_gsea_list(genelist_ranked, genesets)

# save data
export(map(enrich_result_gsea, \(x) as_tibble(x@result)), 
       str_glue("{enrichment_dir}/{Sys.Date()}_enrichment_gsea.xlsx"))
# save plot
n_col <- length(enrich_result_gsea) #if_else( > 1, 2, 1)
map(enrich_result_gsea, 
    \(x) plot_enrich_gsea(x@result, facet_by="Category")) %>%
  plot_grid(plotlist = ., 
            labels = names(enrich_result_gsea), 
            ncol = n_col, 
            axis = "tblr", align = "hv", 
            label_size = 8, scale = 0.9) %>% 
  ggsave(str_glue("{enrichment_dir}/{Sys.Date()}_enrichment_gsea.pdf"),
          .,
          width=n_col*6, 
          height=length(genesets)*4, 
          limitsize = F)
```

### visNetwork
```{r}
pvalue_cutoff <- 1e-3
min_cluster_size <- 4
plist <- list()

for(x in names(enrich_result_gsea)) {
  print(str_glue("visNetwork >>> {x}"))
  plist[[x]] <- enrich_result_gsea[[x]]@result %>% 
    filter(pvalue < pvalue_cutoff) %>% 
    aPEAR::enrichmentNetwork(
      simMethod = "jaccard", # c('jaccard', 'cosine', 'cor'),
      clustMethod = "markov", # c('markov', 'hier', 'spectral'),
      clustNameMethod = "pagerank", # c('pagerank', 'hits', 'none'),
      minClusterSize = min_cluster_size, # min number of nodes in a single cluster
      colorBy = "NES", # pvalue, NES
      colorType = "nes", # use log transform on the colorBy column and adjust color range
      nodeSize = "setSize",
      innerCutoff = 0.1, # similarity cutoff for in-cluster nodes
      outerCutoff = 0.5, # similarity cutoff for between-cluster nodes
      pCutoff = -5, # adjust p-value colouring cutoff when using pval
      drawEllipses = F,
      fontSize = 4,
      repelLabels = F,
      plotOnly = T,
      verbose = F)
}

# save plot
plot_grid(plotlist = plist, 
          labels = names(enrich_result_gsea), 
          ncol = 1, 
          axis = "tblr", align = "hv", 
          label_size = 8, scale = 0.9) %>% 
  ggsave(str_glue(
    "{enrichment_dir}/{Sys.Date()}_enrichment_gesa_visNetwork.pdf"),
    .,
    width=10, 
    height=length(enrich_result_gsea)*10, 
    limitsize = F)
```


### compare gseaplot
```{r}
source("~/Documents/GitHub/myScript/R_functions/helper.R")

items <- c(
  "HALLMARK_MYC_TARGETS_V1",
  "HALLMARK_MYC_TARGETS_V2"
  # "HALLMARK_myc_core_plos_one",
  # "HALLMARK_myc_core_nsmb"
)

for(idx in seq_along(enrich_result_gsea)) {
  # library(patchwork)
  subset <- enrich_result_gsea[[idx]]
  print(names(enrich_result_gsea)[idx])
  pdf(str_glue('{enrichment_dir}/{Sys.Date()}_{names(enrich_result_gsea)[idx]}.pdf'), 
      width=5, height=3.5)
  for(item in items) {
    stat <- subset@result[subset@result$ID==item, c("NES","p.adjust")] %>% 
      mutate(NES=round(NES, 2)) %>% 
      as.numeric()
    # scales::label_number_auto()
    npcx <- ifelse(stat[1] > 0, 0.95, 0.05)
    npcy <- ifelse(stat[1] > 0, 0.95, 0.05)
    anno <- format_num_auto(stat) %>%  str_c(c("NES: ","p.adjust: "), ., collapse="\n")
    p <- gseaplot(subset, geneSetID=item, title=item, 
                  color="grey50", color.line="steelblue", by="runningScore") +
      ggpp::annotate("text_npc", label=anno, size=5, npcx=npcx, npcy = npcy) +
      theme(panel.grid=element_blank())
    # p <- gseaplot2(subset, geneSetID=item, title=item)
    print(p)
  }
  while(dev.cur() != 1) dev.off()
}


compare_list <- list(
  c("Kelly.40nM.1h_vs_Kelly.0nM.16", "SHEP.40nM.1h_vs_SHEP.0nM.16h"),
  c("Kelly.40nM.4h_vs_Kelly.0nM.16", "SHEP.40nM.4h_vs_SHEP.0nM.16h"),
  c("Kelly.40nM.16h_vs_Kelly.0nM.1", "SHEP.40nM.16h_vs_SHEP.0nM.16h"),
  c("Kelly.200nM.1h_vs_Kelly.0nM.1", "SHEP.200nM.1h_vs_SHEP.0nM.16h"),
  c("Kelly.200nM.4h_vs_Kelly.0nM.1", "SHEP.200nM.4h_vs_SHEP.0nM.16h"),
  c("Kelly.200nM.16h_vs_Kelly.0nM.", "SHEP.200nM.16h_vs_SHEP.0nM.16")
)

for(compare in compare_list) {
  # subsets <- enrich_result_gsea[compare[1]]
  print(str_c(str_replace_all(compare, "_vs_.*$", ""), collapse="__"))
  f_name <- str_glue(
    '{enrichment_dir}/{Sys.Date()}_compare_{str_c(str_replace_all(compare, "_vs_.*$", ""), collapse="__")}.pdf')
  
  pdf(f_name, width=4, height=1.5*length(items))
  for(item in items) {
    # item <- "HALLMARK_MYC_TARGETS_V2"
    gsdata <-  enrich_result_gsea[compare] %>% 
      map(\(x) enrichplot:::gsInfo(x, item)) %>% 
      list_rbind(names_to="sample") %>% 
      as_tibble()
    i <- 0
    for (term in unique(gsdata$Description)) {
      idx <- which(gsdata$ymin != 0 & gsdata$Description == term)
      gsdata[idx, "ymin"] <- i
      gsdata[idx, "ymax"] <- i + 1
      i <- i + 1
    }
    
    p1 <- ggplot(gsdata, aes(x, runningScore, group=sample, color=sample)) + 
      geom_hline(yintercept=0, linetype=2, linewidth=1, color="grey50") +
      geom_line(linewidth = 1.2) +
      scale_x_continuous(expand = c(0, 0)) +
      xlab(NULL) + ylab("Running Enrichment Score") + 
      ggtitle(item) +
      theme_classic(8) + 
      theme(panel.grid =element_blank(), 
            legend.position = c(0.8, 0.8), 
            legend.title = element_blank(), 
            legend.background = element_rect(fill = "transparent"))
    p2 <- ggplot(gsdata, aes(x)) + 
      geom_linerange(aes(ymin = ymin, ymax = ymax, color = sample), show.legend=F) + 
      xlab(NULL) + ylab(NULL) +
      scale_x_continuous(expand = c(0,0)) +
      scale_y_continuous(expand = c(0, 0)) +
      facet_wrap(~ sample, ncol=1, strip.position="top") +
      theme_bw(8) +
      theme(panel.grid =element_blank(), 
            strip.text = element_blank(),
            axis.ticks = element_blank(), 
            axis.text = element_blank())
    p2 <- ggrastr::rasterize(p2, dpi=300)
    
    p <- p1 + p2 + plot_layout(ncol=1, heights=c(3, 1))
    
    print(p)
  }
  while(dev.cur() != 1) dev.off()
}

```


## KEGG pathview
```{r}
# save multiple comparison KEGG pathview
if(!dir.exists(file.path(enrichment_dir, "KEGG_pathview"))) 
  dir.create(file.path(enrichment_dir, "KEGG_pathview"))

# setwd (only works in this chunk)
setwd(file.path(enrichment_dir, "KEGG_pathview"))

# construct cross-sample all degs logFC matrix
wald_test_sig_all <- wald_test_sig %>% 
  unlist() %>% unique()
degs_fc_matrix <- map(wald_test, \(x) column_to_rownames(x, "symbol") %>%
                        .[wald_test_sig_all, ] %>%
                        pull(log2FoldChange)) %>%
  bind_cols() %>%
  as.data.frame(row.names = wald_test_sig_all)

print("pathview multi sample ordered as:")
print(str_c(seq_along(colnames(degs_fc_matrix)), 
            colnames(degs_fc_matrix), sep = ": "), sep = "\n")

# show top N KEGG pathways in each comparison
top_kegg <- map(enrich_result_gsea, 
                \(x) dplyr::filter(x@result, Category=="KEGG") %>% 
                  dplyr::slice_min(pvalue, n = 10) %>% 
                  pull(ID)) %>%
  unlist() %>% unique()

kegg_species <- if_else(species == "hs", "hsa", "mmu")
org_db <- if_else(species == "hs", "org.Hs.eg.db", "org.Mm.eg.db")

# pdf
pathview(gene.data = degs_fc_matrix, 
         gene.idtype = "SYMBOL", 
         pathway.id = top_kegg,
         kegg.dir = ".", 
         species = kegg_species, 
         gene.annotpkg = org_db,
         kegg.native = FALSE,
         limit = list(gene = quantile(as.matrix(degs_fc_matrix), 
                                      probs = c(0.02, 0.98))),
         low = list(gene = "#377eb8"), 
         mid = list(gene = "white"), 
         high = list(gene = "#e41a1c"))

```



# Sample Based
## ssGSEA with Hallmark
```{r}
# run many algorithms
gsvapar <- gsvaParam(expr_mat_filtered, genesets$Hallmark)
gsva_res <- GSVA::gsva(gsvapar)
pheatmap(
  gsva_res, border_color=NA, scale='none', 
  cluster_cols = F,
  cluster_rows = T,
  fontsize_row = 6,
  fontsize_col = 4,
  filename=str_glue(
    "{enrichment_dir}/{Sys.Date()}_ssGSEA_heatmap_gsva.pdf"),
  width=6, height=7
)
while(dev.cur() != 1) dev.off()
```

## decouple network
```{r}
# pathway
pathway_progeny <-
  # decoupleR::get_progeny(organism, top = 500) %>%
  {if(params$species=="hs") {
    progeny::model_human_full 
  } else {
    progeny::model_mouse_full
  }} %>% 
  as_tibble() %>%
  arrange(pathway, p.value) %>%
  filter(p.value < 0.001) %>% 
  # dplyr::slice_head(n=500, by=pathway) %>%
  dplyr::select(source=pathway, 
                target=gene, 
                mor=weight, 
                p_value=p.value) %>%
  mutate(across(where(is.factor), as.character))
# pathway_progeny <- get_progeny(organism, top=500) %>%
#   rename(mor=weight)

# TF regulon
tf_collectri <- ifelse(params$species=='hs', 'human', 'mouse') %>% 
  decoupleR::get_collectri(split_complexes=F)

```

## decouple analysis
```{r}
# run pathway
pathway_res <- decouple(expr_mat_filtered, pathway_progeny)
pathway_res$statistic %>% unique()

# run TF
tf_res <- decouple(expr_mat_filtered, tf_collectri)


# save data
write_xlsx(
  list(progeny = pathway_res, collectri = tf_res),
  str_glue("{enrichment_dir}/{Sys.Date()}_activity.xlsx"))

```

## decouple visualize
```{r}
pathway_res$statistic %>% unique()
method_use <- "consensus"

# pathway
# barplot
pathway_res %>%
  left_join(sample, by=join_by(condition==sample_name)) %>%
  # average sample score
  summarise(score=mean(score),
            .by=c(statistic, source, group)) %>%
  # filter(statistic=='consensus') %>%
  ggplot(aes(source, score, fill=group)) +
  # geom_jitter() +
  geom_col(width=0.7, color=NA, position=position_dodge2()) +
  facet_wrap(~ statistic, scales='free') +
  labs(x="") +
  theme_bw() +
  theme(axis.text.x=element_text(angle=45, hjust=1),
        legend.position='top',
        panel.grid = element_blank(),
        strip.background = element_rect(fill=NA))
ggsave(str_glue(
  "{enrichment_dir}/{Sys.Date()}_activity_progeny_barplot.pdf"), 
  width=12, height=6)

# heatmap
pathway_res_mat <- pathway_res %>%
  filter(statistic == method_use) %>%
  pivot_wider(id_cols = 'condition', 
              names_from = 'source',
              values_from = 'score') %>%
  column_to_rownames('condition') %>%
  as.matrix() %>% 
  scale() %>% 
  t() %>% 
  # reorder
  .[, sample$sample_name]
# colnames(pathway_res_mat)
# pathway_res_mat[1:5,1:2]
pheatmap(pathway_res_mat, 
         border_color = NA, 
         color=colorRampPalette(c("#377eb8", "#f7f7f7", "#e41a1c"))(100),
         cluster_rows = T,
         cluster_cols = F,
         annotation_col = column_to_rownames(sample[, -2], "sample_name"),
         filename = str_glue(
           "{enrichment_dir}/{Sys.Date()}_activity_progeny_heatmap.pdf"),
         width = ceiling(nrow(pathway_res_mat)/2), 
         height = ceiling(nrow(pathway_res_mat)/3),
         main = str_glue("Heatmap of pathway activity")
         )
while(dev.cur() != 1) dev.off()




# # TF
# # barplot
# tf_res_5 <- tf_res %>%
#   left_join(sample, by=join_by(condition==sample_name)) %>%
#   reframe(score=mean(score), 
#           p_value=mean(p_value, na.rm=T), 
#           .by=c(statistic, source, group)) %>%
#   reframe(var=var(score), 
#           .by=c(statistic, source)) %>%
#   slice_max(order_by=abs(var), n=200, with_ties = F)
# 
# tf_res_25 <- tf_res %>%
#   arrange(source) %>% 
#   left_join(sample, by=join_by(condition==sample_name)) %>%
#   reframe(score=mean(score),
#           p_value=mean(p_value, na.rm=T),
#           .by=c(statistic, source, group)) %>%
#   reframe(var=var(score), .by=c(statistic, source)) %>%
#   slice_max(order_by=abs(var), n=5, by=c(statistic))
# tf_res %>% 
#   left_join(sample, by=join_by(condition==sample_name)) %>%
#   filter(source %in% tf_res_25$source) %>% 
#   reframe(score=mean(score), 
#           p_value=mean(p_value, na.rm=T), 
#           .by=c(statistic, source, group)) %>% 
#   ggplot(aes(source, score, fill=group)) +
#   geom_col(width=0.7, color=NA, position=position_dodge2()) +
#   facet_wrap(~ statistic, scales='free') +
#   labs(x="") +
#   theme_bw() +
#   theme(axis.text.x=element_text(size=6, angle=45, hjust=1),
#         legend.position='top',
#         panel.grid = element_blank(),
#         strip.background = element_rect(fill=NA))
# ggsave(str_glue(
#   "{enrichment_dir}/{Sys.Date()}_activity_collectri_barplot.pdf"),
#   width=12, height=6)

# heatmap
tf_res_100 <- tf_res %>%
  filter(statistic == method_use) %>%
  left_join(sample, by=join_by(condition==sample_name)) %>%
  reframe(score=mean(score), 
          p_value=mean(p_value, na.rm=T), 
          .by=c(statistic, source, group)) %>%
  reframe(var=var(score), 
          .by=c(statistic, source)) %>%
  slice_max(order_by=abs(var), n=200, with_ties = F)
tf_res_mat <- tf_res %>%
  filter(statistic == method_use) %>%
  # filter(source %in% tf_res_100$source) %>%
  pivot_wider(id_cols = 'condition', 
              names_from = 'source',
              values_from = 'score') %>%
  column_to_rownames('condition') %>%
  as.matrix() %>% 
  scale() %>% 
  t() %>% 
  # reorder
  .[, sample$sample_name]
# dim(tf_res_mat)
# tf_res_mat[1:5,1:2]
pheatmap(tf_res_mat, 
         border_color = NA, 
         color=colorRampPalette(c("#377eb8", "#f7f7f7", "#e41a1c"))(100),
         cluster_rows = T,
         cluster_cols = F,
         show_rownames = F,
         fontsize_row = 3,
         annotation_col = column_to_rownames(sample[, -2], "sample_name"),
         filename = str_glue(
           "{enrichment_dir}/{Sys.Date()}_activity_collectri_heatmap.pdf"),
         width = ceiling(nrow(pathway_res_mat)/2), 
         height = ceiling(nrow(pathway_res_mat)/2),
         main = str_glue("Heatmap of TFs activity")
         )
while(dev.cur() != 1) dev.off()

```

## individual pathway
```{r}
# define pathway of interest
poi <- "MAPK"
label_genes <- c("EPHA2","SPRY4","DUSP6","CCND1")
# label_genes <- NULL

# read deg results
wald_test_file <- list.files(deseq2_dir, 
                             wald_test_suffix,
                             full.names = T) %>% 
  str_subset("^[^~]") %>% 
  last()
comparison <- excel_sheets(wald_test_file)
cat(str_glue("{str_c(comparison, collapse='\n')}"))
wald_test_deg_in_poi <- comparison %>% 
  set_names(comparison) %>% 
  map(\(x) read_xlsx(wald_test_file, x) %>% 
        # order by stats first
        arrange(desc(stat), pvalue) %>% 
        filter(baseMean > 500, 
               abs(log2FoldChange) > deg_fc,
               padj < deg_padj) %>% 
        inner_join(
          filter(pathway_progeny, source == poi, p_value < 0.001),
          by=join_by(symbol==target)) %>% 
        {if(is.null(label_genes)) {.} else{
            mutate(., label=if_else(symbol %in% label_genes, symbol, NA))
          } } %>%
        mutate(label=ifelse(mor>1 & abs(stat)>2, symbol, NA)) %>% 
        mutate(type=case_when(mor > 0 & stat > 0 ~ "1",
                              mor > 0 & stat < 0 ~ "2",
                              mor < 0 & stat > 0 ~ "2",
                              mor < 0 & stat < 0 ~ "1"))
      )

# plot each wald test group
wald_test_deg_in_poi %>%
  map(\(x) ggplot(x, aes(x = mor, y = stat, color = type)) + 
        geom_point(size=0.5, alpha=0.7) +
        # scale_colour_manual(values = c("red","royalblue3","grey")) +
        # geom_label_repel(aes(label = ID)) + 
        geom_text_repel(aes(label=label), size=2) +
        labs(x="Weight",y="Wald test statistics") +
        theme_minimal() +
        theme(legend.position = "none") +
        geom_vline(xintercept = 0, linetype = 'dotted') +
        geom_hline(yintercept = 0, linetype = 'dotted')) %>% 
  plot_grid(plotlist = ., labels = names(.), label_size = 8)

ggsave(str_glue(
  "{enrichment_dir}/{Sys.Date()}_activity_{poi}.pdf"), 
  width=9, height=9)

```






